A full approximation storage multilevel method \\ for nonlinear variational inequalities

### 2022-12-07 paper/mcd2.tex
edits though end section 5

\begin{document}
\title[Multilevel constraint decomposition methods]{Multilevel constraint decomposition methods \\ for nonlinear variational inequalities}

@@

We extend the constraint decomposition (CD) method of X.-C.~Tai \cite{Tai2003} by implementing fully-nonlinear corrections, and we apply it to nonlinear variational inequality (VI) problems to which it has not been applied.  In its multilevel form the method has been shown to have optimal complexity for elliptic, linear obstacle problems \cite[Subsection 5.4]{Tai2003}; see also \cite[Theorem 4.6 and Algorithm 4.7]{GraeserKornhuber2009}.  The convergence of this method, for several types of decompositions, was proven for coercive VI problems which arise from minimization of a convex functional over a convex set \cite{Tai2003}.

@@

For the classical obstacle problem with a Laplacian operator, certain multilevel techniques, such as the truncated monotone multigrid method \cite{Kornhuber1994}, are known to improve performance relative to the original multilevel CD method from \cite{Tai2003}; see also \cite{GraeserKornhuber2009}.  These improved methods either track the active set in the discretization or modify the nodal basis functions, and in this sense they are discrete algorithms, while by contrast the CD method applies essentially at the level of the continuum problem; see Section \ref{sec:cd}.  Acceleration of our NMCD method via active-set and/or basis-level manipulations represents a potential extension of the method here, and is a topic for future research.

@@

Suppose there are $m<\infty$ closed subspaces $\cV_i \subset \cV$ so that the sum
\begin{equation}
\cV = \sum_{i=0}^{m-1} \cV_i \label{eq:subspacedecomp}
\end{equation}
holds in the sense that if $w \in \cV$ then there exist $w_i \in \cV_i$ so that $w = \sum_i w_i$.  Equation \eqref{eq:subspacedecomp} is called a \emph{subspace decomposition} of $\cV$ \cite{Xu1992}.

@@

Finally, for each $\cK_i$ suppose there are bounded, generally-nonlinear \emph{decomposition operators}\footnote{Denoted ``$R_i$'' in \cite{Tai2003}.  Here we avoid a notational conflict with multilevel transfer operators (Section \ref{sec:femultilevel}).} $\Pi_i : \cK \to \cK_i$ such that if $v \in \cK$ then
\begin{equation}
v = \sum_{i=0}^{m-1} \Pi_i v.  \label{eq:constraintrestrictionsum}
\end{equation}
Clearly \eqref{eq:constraintrestrictionsum} implies sense \emph{(i)} for \eqref{eq:constraintdecomp}. A \emph{constraint decomposition} (CD) of $\cK$ is a choice of $\cV_i,\cK_i,\Pi_i$ satisfying \eqref{eq:subspacedecomp}--\eqref{eq:constraintrestrictionsum} \cite{Tai2003}.  Figure \ref{fig:cartoon} suggests how a low-dimensional example might look.

@@

In Sections \ref{sec:femultilevel} and \ref{sec:vcycle} we will introduce discretizations and practical algorithms for finite-dimensional VI problems.  However, one can observe that the CD concept applies at the level of the continuum problem, and the following two obstacle problem examples are illustrations.  The first is an overlapping domain decomposition.

@@

From now on we are concerned with the approximate solution of certain nonlinear VI problems over 2D computational domains.  Suppose $\Omega \subset \RR^2$ is a (open) bounded polygon and that $\mathcal{V}=W^{1,p}(\Omega)$ is a Sobolev space of once-differentiable functions.  Though the free boundaries are defined by inequality constraints (below), fixed boundary conditions may be Dirichlet or mixed.  Suppose that the domain boundary is split into disjoint, measurable sets, $\partial\Omega = \partial_D \Omega \cup \partial_N \Omega$, with $\partial_D \Omega$ of positive measure for simplicity, to permit well-posedness \cite{Evans2010}, and $\partial_N \Omega$ possibly empty.  Suppose that the provided Dirichlet data $g_D:\partial_D \Omega \to \RR$ is sufficiently-regular to permit well-posedness of the VI problem stated below.

We restrict to the case of \emph{box constraints} \cite{BensonMunson2006,FerrisPang1997}, equivalently \emph{bi-obstacle problems}.  The \emph{lower obstacle} $\underline{\gamma}$ and the \emph{upper obstacle} $\overline{\gamma}$ are given measurable functions on $\overline{\Omega}$ which take their values in the extended real line $\tilde\RR = [-\infty,+\infty]$.  Specifically suppose $\underline{\gamma} : \overline{\Omega} \to [-\infty,+\infty)$ and $\overline{\gamma} : \overline{\Omega} \to (-\infty,+\infty]$ are measurable, satisfying $\underline{\gamma} \le \overline{\gamma}$ on $\overline{\Omega}$ and $\underline{\gamma} \le g_D \le \overline{\gamma}$ on $\partial_D \Omega$.

@@

We will solve the same VI as in abstract problem \eqref{eq:vi}.  Given continuous $f:\cK \to \cV'$, and $\ell \in \cV'$, we assume that the following VI is well-posed for $u^*\in \cK$:
\begin{equation}
\ip{f(u^*)}{v-u^*} \ge \ip{\ell}{v-u^*} \qquad \text{for all } v\in \cK. \label{eq:boxdirichletvi}
\end{equation}
Note that $v-u^* \in W_0^{1,p}(\Omega)$, that is, the test functions used in \eqref{eq:boxdirichletvi} have zero values on the Dirichlet boundary $\partial_D\Omega$.

Problem \eqref{eq:boxdirichletvi} is both a free-boundary problem, because of the inequality constraints in defining $\cK$, and a (mixed) fixed-boundary value problem.   Dirichlet conditions $u=g_D$ apply over $\partial_D\Omega$ and Neumann conditions over $\partial_N \Omega$.  The latter are determined by the formula for $f$; for nonhomogeneous Neumann conditions, the nonlinear operator $f:\cK\to\cV'$ will include an integration over $\partial_N\Omega$.

We now construct a multilevel finite element (FE) approximation to problem \eqref{eq:boxdirichletvi} using either triangular or quadrilateral meshes over $\Omega$, and $P_1$ or $Q_1$ elements \cite{Elmanetal2014}, respectively.  The solution on the finest-level mesh, in a hierarchy of meshes constructed by refinement of a coarsest mesh, is the target.   This hierarchy also generates a multilevel CD of $\cK$.  The subsets $\cK_i$ in this CD are designed to allow multigrid-type coarser-level corrections (next Section) in a manner which preserves admissibility, while not introducing unnecessary high frequencies into the coarser-level problems, especially at the free boundary.

The mesh hierarchy and FE spaces are constructed by nested refinement in the following standard manner \cite{Elmanetal2014}.  Suppose $\mathcal{T}^0$ is the \emph{coarsest mesh}, a finite set of non-overlapping, nondegenerate elements (triangles or quadrilaterals) with union equal to $\overline{\Omega}$ and maximum edge length (characteristic size) $h_0>0$.  For $j=1,\dots,J$, let $\mathcal{T}^j$ be the uniform refinement of $\mathcal{T}^{j-1}$ by connecting the midpoints of edges.  For triangles and rectangles, each $\mathcal{T}^{j-1}$ element becomes four similar elements in $\mathcal{T}^j$ with size $h_j = 2^{-j} h_0$.\footnote{For arbitrary triangles, and for rectangles, the refinement generates similar elements and $h_j=2^{-j}h_0$ is the exact maximum edge length in $\mathcal{T}^j$.  For general quadrilaterals, however, similarity is lost and maximum edge lengths are variable; cf.~\cite{Zhang2004}.  Only rectangular quadrilaterals are used in the Results section.}  We call $J\ge 0$ the \emph{depth} of the hierarchy and $\mathcal{T}^J$ the \emph{finest mesh}.  Let $m_j$ be the number of nodes in $\mathcal{T}^j$, the \emph{degrees of freedom} at the $j$th level, and denote $h=h_J=2^{-J} h_0$ and $m=m_J$ for the finest level.  Let $\mathcal{V}^j$ be the continuous $P_1$ or $Q_1$ FE space \cite{Elmanetal2014} over $\mathcal{T}^j$, with arbitrary boundary values, and observe the following vector space nesting:
\begin{equation}
\mathcal{V}^0 \subset \mathcal{V}^1 \subset \dots \subset \mathcal{V}^J \subset \mathcal{V}=W^{1,p}(\Omega).  \label{eq:fe:nestedspaces}
\end{equation}

Use of higher-order Lagrange elements $P_k$ or $Q_k$, $k\ge 2$, would be a ``variational crime'' \cite[Chapter 10]{BrennerScott2007}, specific to inequality constraints, in the upcoming material.  For this reason we restrict to $k = 1$.  Specifically, our FE construction of constraint sets will require \emph{nodal monotonicity}, namely that for all FE functions $w,z \in \mathcal{V}^j$,
\begin{equation}
\bw \ge \bz \quad \implies \quad w \ge z \label{eq:nodalmonotonicity}
\end{equation}
where $\bw,\bz$ denote vectors of nodal values.  For $k\ge 2$ there are hat functions in $P_k$ and $Q_k$, i.e.~functions with nodal value one at a single node, otherwise zero, which take on negative values \cite[Figure 1.7]{Elmanetal2014}.  It is easy to show that \eqref{eq:nodalmonotonicity} is equivalent to the non-existence of sign-changing hat functions, so \eqref{eq:nodalmonotonicity} implies $k\le 1$.

To accomodate the fixed-boundary conditions, let $g_D^J \in \mathcal{V}^J$, on the finest-level mesh, satisfy $g_D^J = g_D$ on $\partial_D \Omega$.\footnote{As is standard in conforming FE methods, this restricts the original Dirichlet boundary values $g_D$ to be a piecewise-linear function \cite{Elmanetal2014}.}  To allow box constraints, let $\tilde{\mathcal{V}}^j$ denote the set of all functions on the nodes of $\mathcal{T}^j$ with values in the extended real line $\tilde{\RR} = [-\infty,+\infty]$.  (Note that $\tilde{\mathcal{V}}^j$ functions are not in $\mathcal{V}^j$.  They are not even defined everywhere on $\Omega$, but only on nodes.)  Suppose that the \emph{lower} and \emph{upper obstacles} $\underline{\gamma}^J, \overline{\gamma}^J \in \tilde{\mathcal{V}}^J$, defined on the nodes of the finest-level mesh $\mathcal{T}^J$, satisfy
\begin{equation}
\underline{\gamma}^J \le \overline{\gamma}^J, \qquad \underline{\gamma}^J < +\infty, \qquad -\infty < \overline{\gamma}^J. \label{eq:fe:boxconstraintrequirements}
\end{equation}
Where $\underline{\gamma}^J=-\infty$ or $\overline{\gamma}^J=+\infty$ then there is no lower or upper constraint, respectively.  Finally assume that $\underline{\gamma}^J \le g_D^J \le \overline{\gamma}^J$ on nodes of $\mathcal{T}^J$ which lie in $\partial_D \Omega$.

Define
\begin{equation}
\mathcal{K}^J = \left\{w\,:\,\underline{\gamma}^J \le w \le \overline{\gamma}^J, \, \text{ and } \, w|_{\partial_D\Omega} = g_D|_{\partial_D\Omega}\right\} \subset \mathcal{V}^J, \label{eq:fe:fineconstraintset}
\end{equation}
as the finest-level \emph{constraint set} of \emph{admissible} FE functions.  This is a nonempty, closed, and convex subset of $\mathcal{V}^J$.  As $\mathcal{V}^J$ functions never take values $\pm\infty$, inequality in \eqref{eq:fe:fineconstraintset} is strict where $\underline{\gamma}^J, \overline{\gamma}^J = \pm \infty$.

Let $f^J:\mathcal{K}^J \to (\mathcal{V}^J)'$ be a discretization of $f$ in \eqref{eq:boxdirichletvi}.  By inclusion \eqref{eq:fe:nestedspaces} the functional $\ell$ is already defined and continuous over $\mathcal{V}^J$, but we denote $\ell^J \in (\mathcal{V}^J)'$ for its FE representation.  We seek the FE solution $u^J \in \mathcal{K}^J$ of the finest-level, finite-dimensional, nonlinear VI
\begin{equation}
\ip{f^J(u^J)}{v-u^J} \ge \ip{\ell^J}{v-u^J} \qquad \text{for all } v\in \cK^J. \label{eq:fe:vi}
\end{equation}
As is true for \eqref{eq:vi}, if $f^J$ is continuous and $p$-coercive on $\mathcal{K}^J$, in the sense that \eqref{eq:pcoercive} holds for all $u,v \in \mathcal{K}^J$, then \eqref{eq:fe:vi} is well-posed.

While \eqref{eq:fe:vi} approximates \eqref{eq:boxdirichletvi}, it may not technically be conforming for that VI.  That is, even if quadrature and other implementation details for $f^J$ are ignored, and despite the vector space inclusion $\mathcal{V}^J \subset \mathcal{V}$, we still do not require $\mathcal{K}^J \subset \mathcal{K}$.  The continuum constraint set $\mathcal{K}$ remains abstract in those applications where only the nodal obstacles $\underline{\gamma}^J, \overline{\gamma}^J$ are defined.  In fact, even if smooth continuum obstacles exist, admissibility with respect to their FE interpolants will generally not imply continuum admissiblity.\footnote{Such caveats apply in geophysical situations where the obstacle is given by meshed elevation data, the continuum limit of which is generally unknown \cite{Bueler2016}.}

Finite-dimensional, box-constrained VI problem \eqref{eq:fe:vi} is equivalent to a \emph{mixed complementarity problem} (MCP) \cite{FacchineiPang2003,FerrisPang1997}.  Stating this strong-form problem may be clarifying for some readers.  Denoting $x_p$ for nodes of the finest-level mesh $\mathcal{T}^J$, let $X=\big\{w\in\cV^J\,:\,w(x_p)=g_D^J(x_p) \text{ if } x_p\in \partial_D\Omega\big\}$, thus $\cK^J = \left\{\underline{\gamma}^J \le w \le \overline{\gamma}^J\right\} \cap X$.  If $m_D$ is the number of nodes $x_p$ in the Dirichlet boundary, let us define $I_D$ as the $m_D\times m_J$ matrix which indicates (extracts) Dirichlet nodal values; its transpose $I_D^\top$ injects Dirichlet values as nodes in $(\mathcal{V}^J)'$.  Problem \eqref{eq:fe:vi} could be stated as VI$\left(f^J(\cdot)-\ell^J,\left[\underline{\gamma}^J,\overline{\gamma}^J\right] \cap X\right)$, for instance in the operations research literature \cite{FerrisPang1997}, and it is equivalent to the equality-constrained MCP
\begin{align}
\underline{\gamma}^J \le u^J \le \overline{\gamma}^J \quad &\perp \quad f^J(u^J) - \ell^J - I_D^\top \lambda \label{eq:fe:mcp} \\
I_D u^J &= g_D^J\big|_{\partial_D\Omega}. \notag
\end{align}
The Lagrange multipliers $\lambda \in \RR^{m_D}$ are free (unconstrained).  Let $\psi_p(x)$ denote the hat function for a node $x_p \in \mathcal{T}^J$.  MCP \eqref{eq:fe:mcp} says that at each node exactly\footnote{The stated conditions are correct anywhere $\underline{\gamma}^J(x_p) < \overline{\gamma}^J(x_p)$.  Where the box constraints meet (pinch) at $x_p \notin \partial_D\Omega$ then $\underline{\gamma}^J(x_p)=u^J(x_p)=\overline{\gamma}^J(x_p)$ holds and $\ip{f^J(u^J)-\ell^J}{\psi_p}$ can have any value.} one of the following conditions holds for the solution $u^J$:
\begin{itemize}
\item $x_p \notin \partial_D\Omega$ and $\underline{\gamma}^J(x_p)<u^J(x_p)<\overline{\gamma}^J(x_p)$ and $\ip{f^J(u^J)-\ell^J}{\psi_p} = 0$, or
\item $x_p \notin \partial_D\Omega$ and $u^J(x_p)=\underline{\gamma}^J(x_p)$ and $\ip{f^J(u^J)-\ell^J}{\psi_p} \ge 0$, or
\item $x_p \notin \partial_D\Omega$ and $u^J(x_p)=\overline{\gamma}^J(x_p)$ and $\ip{f^J(u^J)-\ell^J}{\psi_p} \le 0$, or
\item $x_p \in \partial_D\Omega$ and $u^J(x_p)=g_D^J(x_p)$.
\end{itemize}
These cases might be labeled \emph{inactive}, \emph{lower active}, \emph{upper active}, and \emph{boundary}, respectively.  Where case \emph{boundary} applies, $\lambda_p=\ip{f^J(u^J)-\ell^J}{\psi_p}\in\RR$ has any value.

>>>>

\begin{document}
\title[FAS for nonlinear variational inequalities]{A full approximation storage multilevel method \\ for nonlinear variational inequalities}

@@

We extend the constraint decomposition (CD) method of Tai \cite{Tai2003} by implementing fully-nonlinear corrections, and we apply it to a new set of nonlinear variational inequality (VI) problems.  The method has been shown to have almost grid-independent error bounds, and near-optimal complexity, for elliptic, linear obstacle problems \cite[Subsection 5.4]{Tai2003}; see also \cite[Theorem 4.6 and Algorithm 4.7]{GraeserKornhuber2009}.  The convergence of this method, for several types of decompositions, was proven for coercive VI problems which arise from minimization of a convex functional over a convex set \cite{Tai2003}.

@@

For the classical obstacle problem with a Laplacian operator, certain multilevel techniques, such as the truncated monotone multigrid method \cite{Kornhuber1994}, are known to improve performance relative to the original multilevel CD method from \cite{Tai2003}; see also \cite{GraeserKornhuber2009}.  These improved methods either track the active set in the discretization or modify the nodal basis functions, and in this sense they are discrete algorithms.  By contrast the CD method applies essentially at the level of the continuum problem; see Section \ref{sec:cd}.  Acceleration of our NMCD method via active-set and/or basis-level manipulations represents a potential extension of the method here, and is a topic for future research.

@@

If there are $m<\infty$ closed subspaces $\cV_i \subset \cV$ so that the sum
\begin{equation}
\cV = \sum_{i=0}^{m-1} \cV_i \label{eq:subspacedecomp}
\end{equation}
holds in the sense that for $w \in \cV$ there exist $w_i \in \cV_i$ so that $w = \sum_i w_i$ then we say \eqref{eq:subspacedecomp} is a \emph{subspace decomposition} of $\cV$ \cite{Xu1992}.

@@

Finally, for each $\cK_i$ suppose there are bounded, generally-nonlinear \emph{decomposition operators}\footnote{Denoted ``$R_i$'' in \cite{Tai2003}.  Here we avoid a notational conflict with multilevel transfer operators (Section \ref{sec:transfers}).} $\Pi_i : \cK \to \cK_i$ such that if $v \in \cK$ then
\begin{equation}
v = \sum_{i=0}^{m-1} \Pi_i v.  \label{eq:constraintrestrictionsum}
\end{equation}
Clearly \eqref{eq:constraintrestrictionsum} implies sense \emph{(i)} for \eqref{eq:constraintdecomp}. A \emph{constraint decomposition} (CD) of $\cK$ is a choice of $\cV_i,\cK_i,\Pi_i$ satisfying \eqref{eq:subspacedecomp}--\eqref{eq:constraintrestrictionsum} \cite{Tai2003}.  Figure \ref{fig:cartoon} suggests how a low-dimensional example might look.

@@

In Sections \ref{sec:femultilevel}--\ref{sec:smoothers} we will introduce multilevel CD discretizations and algorithms for finite-dimensional VI problems.  However, one can observe here that the CD concept applies at the level of the continuum problem.  The following two obstacle problem examples are illustrations; the first is an overlapping domain decomposition.

@@

From now on we are concerned with the approximate solution of certain nonlinear VI problems over 2D computational domains.  Suppose $\Omega \subset \RR^2$ is a (open) bounded polygon and that $\mathcal{V}=W^{1,p}(\Omega)$ is a Sobolev space of once-differentiable functions.  In addition to the free boundaries defined by inequality constraints (below), fixed boundary conditions may be Dirichlet or mixed.  Suppose that the domain boundary is split into disjoint, measurable sets, $\partial\Omega = \partial_D \Omega \cup \partial_N \Omega$, with $\partial_D \Omega$ of positive measure for simplicity, to permit well-posedness \cite{Evans2010}, and $\partial_N \Omega$ possibly empty.  Suppose that the provided Dirichlet data $g_D:\partial_D \Omega \to \RR$ is sufficiently-regular so that the VI problem \eqref{eq:boxdirichletvi} below is well posed.

We restrict to the case of \emph{box constraints} \cite{BensonMunson2006,FerrisPang1997}, equivalently \emph{bi-obstacle problems}.  The \emph{lower obstacle} $\underline{\gamma}$ and the \emph{upper obstacle} $\overline{\gamma}$ are given measurable functions on the closure $\overline{\Omega}$ which take their values in the extended real line $\tilde\RR = [-\infty,+\infty]$.  Specifically suppose $\underline{\gamma} : \overline{\Omega} \to [-\infty,+\infty)$ and $\overline{\gamma} : \overline{\Omega} \to (-\infty,+\infty]$ are measurable, satisfying $\underline{\gamma} \le \overline{\gamma}$ on $\overline{\Omega}$ and $\underline{\gamma} \le g_D \le \overline{\gamma}$ on $\partial_D \Omega$.

@@

We will solve the same VI as in abstract problem \eqref{eq:vi}.  Given continuous $f:\cK \to \cV'$, and $\ell \in \cV'$, we assume that the following VI is well-posed for $u^*\in \cK$:
\begin{equation}
\ip{f(u^*)}{v-u^*} \ge \ip{\ell}{v-u^*} \qquad \text{for all } v\in \cK. \label{eq:boxdirichletvi}
\end{equation}
Note that $v-u^* \in W_0^{1,p}(\Omega)$, that is, test functions in \eqref{eq:boxdirichletvi} have zero values on the Dirichlet boundary $\partial_D\Omega$.

Problem \eqref{eq:boxdirichletvi} is both a free-boundary problem, because of the inequality constraints in defining $\cK$, and a fixed-boundary value problem.   Dirichlet conditions $u=g_D$ apply over $\partial_D\Omega$ and Neumann conditions over $\partial_N \Omega$.  The latter are determined by the formula for $f$; for nonhomogeneous Neumann conditions, the nonlinear operator $f:\cK\to\cV'$ must include an integration over $\partial_N\Omega$.

We now construct a multilevel finite element (FE) approximation to problem \eqref{eq:boxdirichletvi} using either triangular or quadrilateral meshes over $\Omega$, and $P_1$ or $Q_1$ elements \cite{Elmanetal2014}, respectively.  A hierarchy of meshes will be constructed by refinement of a coarsest mesh.  A solution on the finest-level mesh is the goal.  This hierarchy also allows the creation of a multilevel CD of $\cK$.  The subsets $\cK_i$ in this CD are designed to allow multigrid-type coarser-level corrections (next Section) in a manner which preserves admissibility, while not introducing unnecessary high frequencies into the coarser-level problems, especially at the free boundary.

The mesh hierarchy and FE spaces are constructed by nested refinement in the following standard manner \cite{Elmanetal2014}.  Suppose $\mathcal{T}^0$ is the \emph{coarsest mesh}, a finite set of non-overlapping, nondegenerate elements (triangles or quadrilaterals) with union equal to $\overline{\Omega}$ and maximum edge length (characteristic size) $h_0>0$.  For $j=1,\dots,J$, let $\mathcal{T}^j$ be the uniform refinement of $\mathcal{T}^{j-1}$ by connecting the midpoints of edges.  For triangles and rectangles, each $\mathcal{T}^{j-1}$ element becomes four similar elements in $\mathcal{T}^j$ with size $h_j = 2^{-j} h_0$.\footnote{For arbitrary triangles, and for rectangles, edge midpoint refinement generates similar elements and $h_j=2^{-j}h_0$ is the exact maximum edge length in $\mathcal{T}^j$.  For general quadrilaterals, however, similarity is lost and maximum edge lengths are variable; cf.~\cite{Zhang2004}.  Only rectangular quadrilaterals are used in the Results section.}  We call $J\ge 0$ the \emph{depth} of the hierarchy and $\mathcal{T}^J$ the \emph{finest-level mesh}.  Let $m_j$ be the number of nodes in $\mathcal{T}^j$, the \emph{degrees of freedom} at the $j$th level, and denote $h=h_J=2^{-J} h_0$ and $m=m_J$ for the finest level.

Let $\mathcal{V}^j$ be the continuous $P_1$ or $Q_1$ FE space \cite{Elmanetal2014} over $\mathcal{T}^j$, with arbitrary boundary values, and observe the following vector space nesting:
\begin{equation}
\mathcal{V}^0 \subset \mathcal{V}^1 \subset \dots \subset \mathcal{V}^J \subset \mathcal{V}=W^{1,p}(\Omega).  \label{eq:fe:nestedspaces}
\end{equation}

Use of higher-order Lagrange elements $P_k$ or $Q_k$, $k\ge 2$, would imply a ``variational crime'' \cite[Chapter 10]{BrennerScott2007} special to inequality constraints.  Namely, our FE construction of constraint sets will require \emph{nodal monotonicity},
\begin{equation}
\bw \ge \bz \quad \implies \quad w \ge z \label{eq:nodalmonotonicity}
\end{equation}
for all FE functions $w,z \in \mathcal{V}^j$, where $\bw,\bz$ denote vectors of nodal values.  For $k\ge 2$ there are hat functions in $P_k$ and $Q_k$, i.e.~functions with nodal value one at a single node, otherwise zero, which take on negative values \cite[Figure 1.7]{Elmanetal2014}.  It is easy to show that \eqref{eq:nodalmonotonicity} is equivalent to the non-existence of sign-changing hat functions, so \eqref{eq:nodalmonotonicity} implies $k\le 1$.  For this reason we restrict to $k = 1$.

Let $g_D^J \in \mathcal{V}^J$, on the finest-level mesh, satisfy $g_D^J = g_D$ on $\partial_D \Omega$.  Typically of conforming FE methods, this restricts the Dirichlet boundary function $g_D$ to be a piecewise-linear \cite{Elmanetal2014}.

To allow box constraints, let $\tilde{\mathcal{V}}^j$ denote the set of all functions on the nodes of $\mathcal{T}^j$ with values in the extended real line $\tilde{\RR} = [-\infty,+\infty]$.  (Note that $\tilde{\mathcal{V}}^j$ functions are not in $\mathcal{V}^j$.  They are not defined everywhere on $\Omega$, but only on nodes.)  Suppose that the \emph{lower} and \emph{upper obstacles} $\underline{\gamma}^J, \overline{\gamma}^J \in \tilde{\mathcal{V}}^J$, for nodes of the finest-level mesh $\mathcal{T}^J$, satisfy
\begin{equation}
\underline{\gamma}^J \le \overline{\gamma}^J, \qquad \underline{\gamma}^J < +\infty, \qquad -\infty < \overline{\gamma}^J. \label{eq:fe:boxconstraintrequirements}
\end{equation}
Where $\underline{\gamma}^J=-\infty$ or $\overline{\gamma}^J=+\infty$ then there is no lower or upper constraint, respectively.  Finally assume that $\underline{\gamma}^J \le g_D^J \le \overline{\gamma}^J$ on nodes of $\mathcal{T}^J$ which lie in $\partial_D \Omega$.

Define
\begin{equation}
\mathcal{K}^J = \left\{w\,:\,\underline{\gamma}^J \le w \le \overline{\gamma}^J, \, \text{ and } \, w|_{\partial_D\Omega} = g_D|_{\partial_D\Omega}\right\} \subset \mathcal{V}^J, \label{eq:fe:fineconstraintset}
\end{equation}
as the finest-level \emph{constraint set} of admissible FE functions.  This is a nonempty, closed, and convex subset of $\mathcal{V}^J$.  As $\mathcal{V}^J$ functions never take values $\pm\infty$, inequality in \eqref{eq:fe:fineconstraintset} is strict where $\underline{\gamma}^J, \overline{\gamma}^J = \pm \infty$.

Let $f^J:\mathcal{K}^J \to (\mathcal{V}^J)'$ be a discretization of $f$ in \eqref{eq:boxdirichletvi}.  By inclusion \eqref{eq:fe:nestedspaces} the functional $\ell$ is already defined and continuous over $\mathcal{V}^J$, but we denote $\ell^J \in (\mathcal{V}^J)'$.  We seek the FE solution $u^J \in \mathcal{K}^J$ of the finest-level, finite-dimensional, nonlinear VI
\begin{equation}
\ip{f^J(u^J)}{v-u^J} \ge \ip{\ell^J}{v-u^J} \qquad \text{for all } v\in \cK^J. \label{eq:fe:vi}
\end{equation}
As is true for \eqref{eq:vi}, if $f^J$ is continuous and $p$-coercive on $\mathcal{K}^J$, in the sense that \eqref{eq:pcoercive} holds for all $u,v \in \mathcal{K}^J$, then \eqref{eq:fe:vi} is well-posed.

While \eqref{eq:fe:vi} approximates \eqref{eq:boxdirichletvi}, it may not technically be conforming for that VI.  That is, even if quadrature and other implementation details for $f^J$ are ignored, and despite the vector space inclusion $\mathcal{V}^J \subset \mathcal{V}$, we do not require $\mathcal{K}^J \subset \mathcal{K}$.  Even if smooth continuum obstacles exist, admissibility with respect to their FE interpolants will generally not imply continuum admissiblity.  Such caveats apply in geophysical situations where only nodal values of the obstacle are known, and given by meshed elevation data.  The continuum limit of nodal obstacles $\underline{\gamma}^J$, $\overline{\gamma}^J$ may be unknown \cite{Bueler2016}.

Finite-dimensional, box-constrained VI problem \eqref{eq:fe:vi} is equivalent to a \emph{mixed complementarity problem} (MCP) \cite{FacchineiPang2003,FerrisPang1997}.  Stating this strong-form problem may be clarifying for some readers.  Denoting $x_p$ for nodes of the finest-level mesh $\mathcal{T}^J$, let $X=\big\{w\in\cV^J\,:\,w(x_p)=g_D^J(x_p) \text{ if } x_p\in \partial_D\Omega\big\}$, thus $\cK^J = \left\{\underline{\gamma}^J \le w \le \overline{\gamma}^J\right\} \cap X$.  If $m_D$ is the number of nodes $x_p$ in the Dirichlet boundary, let us define $I_D$ as the $m_D\times m_J$ matrix which indicates (extracts) Dirichlet nodal values; its transpose $I_D^\top$ injects Dirichlet values as nodes in $(\mathcal{V}^J)'$.  Problem \eqref{eq:fe:vi} must be stated as VI$\left(f^J(\cdot)-\ell^J,\left[\,\underline{\gamma}^J,\overline{\gamma}^J\right] \cap X\right)$, in the operations research literature \cite{FerrisPang1997}, and it is equivalent to the equality-constrained MCP
\begin{align}
\underline{\gamma}^J \le u^J \le \overline{\gamma}^J \quad &\perp \quad f^J(u^J) - \ell^J - I_D^\top \lambda \label{eq:fe:mcp} \\
I_D u^J &= g_D^J\big|_{\partial_D\Omega}. \notag
\end{align}
The Lagrange multipliers $\lambda \in \RR^{m_D}$ are free (unconstrained).  Let $\psi_p(x)$ denote the hat function for a node $x_p \in \mathcal{T}^J$.  MCP \eqref{eq:fe:mcp} says that at each node exactly\footnote{The stated conditions are correct anywhere $\underline{\gamma}^J(x_p) < \overline{\gamma}^J(x_p)$.  Where the box constraints meet (pinch) at $x_p \notin \partial_D\Omega$ then $\underline{\gamma}^J(x_p)=u^J(x_p)=\overline{\gamma}^J(x_p)$ holds.} one of the following conditions holds for the solution $u^J$:
\begin{itemize}
\item $x_p \notin \partial_D\Omega$ and $\underline{\gamma}^J(x_p)<u^J(x_p)<\overline{\gamma}^J(x_p)$ and $\ip{f^J(u^J)-\ell^J}{\psi_p} = 0$, or
\item $x_p \notin \partial_D\Omega$ and $u^J(x_p)=\underline{\gamma}^J(x_p)$ and $\ip{f^J(u^J)-\ell^J}{\psi_p} \ge 0$, or
\item $x_p \notin \partial_D\Omega$ and $u^J(x_p)=\overline{\gamma}^J(x_p)$ and $\ip{f^J(u^J)-\ell^J}{\psi_p} \le 0$, or
\item $x_p \in \partial_D\Omega$ and $u^J(x_p)=g_D^J(x_p)$.
\end{itemize}
These cases might be labeled \emph{inactive}, \emph{lower active}, \emph{upper active}, and \emph{boundary}, respectively.  Where case \emph{boundary} applies, $\lambda_p=\ip{f^J(u^J)-\ell^J}{\psi_p}\in\RR$ has any value.
